\texttt{Pythran} -- C++ for Snakes\\\emph{\small Static Compiler for High Performance}

### 2014-05-04 doc/papers/pydata2014/talk.tex
Update timing: -Ofast helped a lot on Mandelbrot!
Finale version of the slides presented @PyData 2014

%\documentclass{beamer} % use draft mode for speed !
\documentclass[aspectratio=1610]{beamer}
\usepackage{pacman}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amsfonts, wasysym}
\usepackage{xcolor}
\usepackage{listings}
\definecolor{darkgreen}{rgb}{.1,.7,.1}
\definecolor{darkred}{rgb}{.7,.1,.1}
\definecolor{darkblue}{rgb}{.1,.1,.7}
\lstset{language=python, basicstyle=\ttfamily,moredelim=[is][\textcolor{darkblue}]{@}{@},keywordstyle=\textcolor{darkblue},commentstyle=\textcolor{darkgreen},stringstyle=\textcolor{darkred},
        backgroundcolor=\color[gray]{0.9},
                frame=single
}

@@

All timings were performed using OSX 10.9, an i7-3740QM CPU @ 2.70GHz, Python 2.7.6, current Pythran (c7de8b9c), pypy 2.2.1, numba 0.13.1, gcc 4.8, Clang r207887.
\vfill
I am \textbf{not} Pythonista, but I'm interested in performance in general. 
Daily job: driver-level C code, assembly, multi-threaded C++, GPU, ...
\vfill
\uncover<2->{
\begin{minipage}{0.4\textwidth}
\textit{Note: I love to troll, so let's have a beer later and talk
about how awful Python is! ;-)}
\end{minipage}
\hspace{0.09\textwidth}
\begin{minipage}{0.4\textwidth}
\uncover<3->{
    \includegraphics[width=0.9\textwidth]{compiling.png}
}
\end{minipage}
}
\vfill
\uncover<4->{
  By the way this talk is written in Latex and takes more than 10 seconds to
  \textbf{compile}. 
}
\end{frame}

@@

\begin{frame}[fragile]
    \frametitle{Explicit Parallelization}
            \begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
def hyantes(xmin, ymin, xmax, ymax, step, range_, range_x, range_y, t):
  pt = [[0]*range_y for _ in range(range_x)]
  "#omp parallel for"
  for i in xrange(range_x):
    for j in xrange(range_y):
      s = 0
      for k in t:
        tmp = 6368.* math.acos(math.cos(xmin+step*i)*math.cos( k[0] ) *
                               math.cos((ymin+step*j)-k[1]) +
                               math.sin(xmin+step*i)*math.sin(k[0]))
        if tmp < range_:
          s+=k[2] / (1+tmp)
      pt[i][j] = s
  return pt
\end{lstlisting}
\vfill
    \begin{tabular}{|l|r|r|r|}
        \hline
     Tool    &  CPython    &   Pythran     & OpenMP   \\
    \hline
     Timing  &  639.0ms    &   44.8ms       &      11.2ms       \\
    \hline
     Speedup &  $\times$1         &    $\times$14.2      &    $\times$57.7   \\
    \hline
\end{tabular}
\end{frame}

@@

\begin{lstlisting}
def run_julia(cr, ci, N, bound, lim, cutoff):
    julia = np.empty((N, N), np.uint32)
    grid_x = np.linspace(-bound, bound, N)
    t0 = time()
    "#omp parallel for"
    for i, x in enumerate(grid_x):
        for j, y in enumerate(grid_x):
            julia[i,j] = kernel(x, y, cr, ci, lim, cutoff)
    return julia, time() - t0
\end{lstlisting}
From Scipy2013 Cython Tutorial.
\vfill
    \begin{tabular}{|l|r|r|r|r|}
        \hline
        Tool    &  CPython    &  Cython  &  Pythran & +OpenMP   \\
        \hline
        Timing  &  3630ms    &   4.3ms  &   3.71ms  & 1.52ms  \\
        \hline
        Speedup & $\times$1 &  $\times$837 &  $\times$970 & $\times$2368   \\
        \hline
    \end{tabular}

@@

    return image
\end{lstlisting}
\end{minipage}
    \vfill
    \begin{tabular}{|l|r|r|r|r|}
        \hline
     Tool    &  CPython    &   Numba    &   Pythran  \\
        \hline
     Timing  &  8170ms   &    56ms       &   83ms  \\
        \hline
     Speedup & $\times$1 & $\times$145 & $\times$98 \\
    \hline
\end{tabular}
\end{frame}

@@

\begin{frame}{Conclusions}
    \hfill\tiny{\url{http://pythonhosted.org/pythran/}}
    \normalsize
    \begin{center}
        \vfill
        Compiling Python means more than typing and translating\\
        \vfill
        \uncover<2-> {
        What next: 
        \begin{itemize}
          \item<3-> Release the PyData version \scriptsize{(ooops we're late)}
          \item<4-> User module import \scriptsize{(pull request already issued)}
          \item<5-> Set-up an daily performance regression test bot \scriptsize{(ongoing work with codespeed)}
          \item<6-> Re-enable vectorization through boost::simd and nt2 \scriptsize{(soon)}
          \item<7-> More Numpy support, start looking into Scipy
          \item<8-> Polyhedral transformations \scriptsize{(in another life...)}
        \end{itemize}
        }
        \vfill
        % \uncover<3->{
        %     Need a compiler guy / HPC Engineer? I am looking for a job!\\
        % }
        \vfill
    \end{center}
\end{frame}

>>>>

%\documentclass{beamer} % use draft mode for speed !
\documentclass[aspectratio=1610]{beamer}
\usepackage{pacman}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amsfonts, wasysym}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{ulem}
\definecolor{darkgreen}{rgb}{.1,.7,.1}
\definecolor{darkred}{rgb}{.7,.1,.1}
\definecolor{darkblue}{rgb}{.1,.1,.7}
\lstset{language=python, basicstyle=\ttfamily,moredelim=[is][\textcolor{darkblue}]{@}{@},keywordstyle=\textcolor{darkblue},commentstyle=\textcolor{darkgreen},stringstyle=\textcolor{darkred},
        backgroundcolor=\color[gray]{0.9},
                frame=single
}

@@

Timings were performed using OSX 10.9, an i7-3740QM CPU @ 2.70GHz, Python 2.7.6, current Pythran (branch Pydata2014), pypy 2.2.1, numba 0.13.1, gcc 4.8, Clang r207887.
\vfill
I am \textbf{not} Pythonista, but I'm interested in performance in general. 
\\ Daily job: driver-level C code, assembly, multi-threaded C++, GPU, ...
\vfill
\uncover<2->{
\begin{minipage}{0.4\textwidth}
\textit{Note: I love to troll, so let's have a beer later and talk
about how awful Python is! ;-)}
\end{minipage}
\hspace{0.09\textwidth}
\begin{minipage}{0.4\textwidth}
\uncover<3->{
    \includegraphics[width=0.9\textwidth]{compiling.png}
}
\end{minipage}
}
\vfill
\uncover<3->{
\scriptsize{
  By the way this talk is written in Latex and takes more than 10 seconds to
  \textbf{compile}.
}
}
\end{frame}

@@

\begin{frame}[fragile]
    \frametitle{Explicit Parallelization}
            \begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
def hyantes(xmin, ymin, xmax, ymax, step, range_, range_x, range_y, t):
  pt = [[0]*range_y for _ in range(range_x)]
  #omp parallel for
  for i in xrange(range_x):
    for j in xrange(range_y):
      s = 0
      for k in t:
        tmp = 6368.* math.acos(math.cos(xmin+step*i)*math.cos( k[0] ) *
                               math.cos((ymin+step*j)-k[1]) +
                               math.sin(xmin+step*i)*math.sin(k[0]))
        if tmp < range_:
          s+=k[2] / (1+tmp)
      pt[i][j] = s
  return pt
\end{lstlisting}
\vfill
    \begin{tabular}{|l|r|r|r|}
        \hline
     Tool    &  CPython    &   Pythran     & OpenMP   \\
    \hline
     Timing  &  639.0ms    &   44.8ms       &      11.2ms       \\
    \hline
     Speedup &  $\times$1         &    $\times$14.2      &    $\times$57.7   \\
    \hline
\end{tabular}
\end{frame}

@@

\begin{lstlisting}
def run_julia(cr, ci, N, bound, lim, cutoff):
    julia = np.empty((N, N), np.uint32)
    grid_x = np.linspace(-bound, bound, N)
    t0 = time()
    #omp parallel for
    for i, x in enumerate(grid_x):
        for j, y in enumerate(grid_x):
            julia[i,j] = kernel(x, y, cr, ci, lim, cutoff)
    return julia, time() - t0
\end{lstlisting}
From Scipy2013 Cython Tutorial.
\vfill
    \begin{tabular}{|l|r|r|r|r|}
        \hline
        Tool    &  CPython    &  Cython  &  Pythran & +OpenMP   \\
        \hline
        Timing  &  3630ms    &   4.3ms  &   3.71ms  & 1.52ms  \\
        \hline
        Speedup & $\times$1 &  $\times$837 &  $\times$970 & $\times$2368   \\
        \hline
    \end{tabular}

@@

    return image
\end{lstlisting}
\end{minipage}
    \vfill
    \begin{tabular}{|l|r|r|r|r|}
        \hline
     Tool    &  CPython    &   Numba    &   Pythran  \\
        \hline
     Timing  &  8170ms   &    56ms       &   47.2ms  \\
        \hline
     Speedup & $\times$1 & $\times$145 & $\times$173 \\
    \hline
\end{tabular}

\scriptsize{"g++ -Ofast" here}
\end{frame}

@@

\begin{frame}[fragile]{Conclusion}
    \normalsize
        \vfill
        Compiling Python means more than typing and translating\\
        \vfill
        \uncover<2-> {
        \vfill
        What next: 
        \vfill
        \begin{itemize}
          \item<2-> Release the PyData version  
          \only<3>{\scriptsize{(ooops we're late)}} \only<4-> {\scriptsize{(\sout{ooops we're late} I did not specify which year)}}
          \item<4-> User module import \scriptsize{(pull request already issued)}
          \item<5-> Set-up an daily performance regression test bot \scriptsize{(ongoing work with codespeed)}
          \item<6-> Re-enable vectorization through boost::simd and nt2 \scriptsize{(soon)}
          \item<7-> More Numpy support, start looking into Scipy
          \item<8-> Polyhedral transformations \scriptsize{(in another life...)}
        \end{itemize}
        }
        \vfill
        % \uncover<3->{
        %     Need a compiler guy / HPC Engineer? I am looking for a job!\\
        % }
        \vfill
    \hfill\tiny{\url{http://pythonhosted.org/pythran/}}
\end{frame}
